SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-11-15 paper_1.tex
Version 2.3
Typos / misc.

\vspace{1ex}The rest of the paper is organized as follows:

\begin{itemize}
\item Section 2 talks about background work and recaptures the SPP architecture
\item Section 3 discusses enhancements made to the SPP design and the detailed implementation of the perceptron mechanism
\item Section 4 consists of the methodology for evaluating SPP-Perc and exploring the feature space for perceptron learning
\item Section 5 comprises results and evaluation in terms of IPC, MPKI, coverage and DRAM traffic, for a variety of benchmark suites
\item Section 6 is the conclusion and future scope of expansion on this work
\end{itemize}

@@

A related line of work also explores cache replacement - insertion policies and dead block prediction, with an eye towards prefetching. 
\cite{DB_Pred}
\cite{Cache_Burst}
\cite{KPC}
\cite{Harmony}
Nori \textit{et.al.} introduced prefetching with  respect to runtime criticality.
\cite{CATCH}
\newline

\textit{Machine Learning and Prefetching:}
Peled \textit{et.al.} introduce interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
\cite{Semantics}
The biggest challenge here is that it relies on compiler support for getting features to build the context.

@@

et. al.

@@

Perceptron learning for computer architecture design has been around for a while. 
It was first popularized by Jimenez \textit{et. al.} in for doing branch prediction. \cite{Perc_Branch} 
In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. 
The retrieved weights would then be multiplied with the history of branch outcomes stored in the History Register (which is a feature) in a dot product fashion. 
The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the prediction is made as \textit{true}, or branch taken.
Else if the sum is < 0, the prediction is made as \textit{false}.

@@

This mechanism allows us to solve the classical problem of exploiting `lost opportunity'. 
In prior perceptron based implementations (and in general, prefetching algorithms), there is usually no way of knowing the result of not prefetch a particular line.
Our two-step implementation allows us to overcome that issue in a simple manner.

@@

Here we discuss the various features which help correlate the prefetching decision with the program behaviour
\begin{itemize}
\item \textbf{Base address} of the demand access which triggered the prefetch. 
Since prefetch especially tend to be triggered from the L2 demand access, they tend to be correlated with the triggering address.

@@

We performed cross-validation of our SPP-Perc model using SPEC 2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.
For SPEC 2006, we developed 94 simpoints spread across the 29 applications.
For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.
The traces include 4 multi-core applications with 6 distinct phases per application. 
In total, we used 285 traces representing workloads across 53 applications.

\textit{Single-core performance} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics. 
We report the performance as speedup over baseline \textit{i.e.}, no prefetching. 
We take the Instruction Per Cycle (IPC) of the prefetcher and divide it with the IPC of baseline. 
The final number reported is the geometric mean of the speed-up attained on individual traces.

\textit{Multi-core performance} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
For 4-core workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. 
Each CPU keeps executing its workload till the last CPU completes 1 Billion instructions after warm-up. 
For collecting IPC and other data, only the first billion instructions are considered as the ``region of interest''.

>>>>

The rest of the paper is organized as follows.
Section 2 talks about background work and recaptures the SPP architecture
Section 3 discusses the detailed implementation of the perceptron mechanism and the features used.
Section 4 consists of the evaluation methodology and explores the feature space for perceptron learning.
Section 5 is the performance evaluation of our prefetcher and Section 6 concludes the paper.

@@

A related line of work also explores cache replacement - insertion policies and dead block prediction, with an eye towards prefetching. 
\cite{DB_Pred}
\cite{Cache_Burst}
\cite{KPC}
\cite{Harmony}
Nori \textit{et.al.} introduced prefetching with  respect to runtime criticality.
\cite{CATCH}
\newline
\newline
\textit{Machine Learning and Prefetching:}
Peled \textit{et.al.} introduce interesting ideas for on-line Reinforcement Learning and dynamically scaling the magnitude of feedback given to the baseline prefetcher.
\cite{Semantics}
The biggest challenge here is that it relies on compiler support for getting features to build the context.

@@

@@

Perceptron learning for computer architecture design has been around for a while. 
It was first popularized by Jimenez \textit{et. al.} for doing branch prediction. \cite{Perc_Branch} 
In it's original implementation, the Program Counter of the instruction looking for a branch prediction would index into a given table of perceptron weights. 
The retrieved weights would then be multiplied with the history of branch outcomes stored in the History Register (which is a feature) in a dot product fashion. 
The sum obtained is thresholded at 0 \textit{i.e.}, if the prediction y\textsubscript{out} >= 0, the prediction is made as \textit{true}, or branch taken.
Else if the sum is < 0, the prediction is made as \textit{false}.

@@

This mechanism allows us to solve the classical problem of exploiting `lost opportunity'. 
In prior perceptron based implementations (and in general, prefetching algorithms), there is usually no way of knowing the result of not prefetch a particular line.
Our two-step SPP-Perc allows us to overcome that issue in a simple manner.

@@

Here we discuss the various features which help correlate the prefetching decision with the program behaviour
\begin{itemize}
\item \textbf{Base address} of the demand access which triggered the prefetch. 
Since prefetch are triggered from the L2 demand access, they tend to be correlated with the triggering address.

@@

We performed cross-validation of our SPP-Perc model using SPEC 2006\cite{SPEC2006} and CloudSuite\cite{CloudSuite} benchmarks.
For SPEC 2006, we developed 94 simpoints spread across the 29 applications.
For CloudSuite, we used the traces made available for the 2nd Cache Replacement Competition (CRC-2)\cite{CRC_2}.
The traces include 4 multi-core applications with 6 distinct phases per application. 
In total, we used 285 traces representing workloads across 53 applications. 
\newline
\newline
\textit{Single-core performance:} For single-core simulations, we use 200 Million instructions to warm-up the microarchitectural structures and next 1 Billion instructions to do detailed simulations and collect run-time statistics. 
We report the performance as speedup over baseline \textit{i.e.}, no prefetching. 
We take the Instruction Per Cycle (IPC) of the prefetcher and divide it with the IPC of baseline. 
The final number reported is the geometric mean of the speed-up attained on individual traces.
\newline
\newline
\textit{Multi-core performance:} For multi-application workloads, we generate 100 completely random mixes and another 100 mixes from the memory intensive subset of SPEC 2017.
For 4-core workload, 200 Million instructions are used for warm-up and additional 1 Billion instruction simulated for collecting statistics. 
Each CPU keeps executing its workload till the last CPU completes 1 Billion instructions after warm-up. 
For collecting IPC and other data, only the first billion instructions are considered as the ``region of interest''.
